### 2024-01-02 latex/tikz/frame_structure.tex

\begin{tikzpicture}[
  , thick
  , inner sep = 0
  , outer sep = 0
  , n/.style = {
    , draw
    , fill = #1!15
    , minimum height = 12pt
  }
  , node distance = -.5\pgflinewidth
  , font = \small\sffamily
  , dl/.style = {
    , semithick
    , densely dotted
  }
]
  \node [n = myblued, minimum width = 18pt] (TRN-1) {$\mathbf{D}$};
  \node [n = myblued, minimum width = 18pt, right = of TRN-1] (TRN-2) {$\mathbf{D}$};
  \node [n = myblued, minimum width = 18pt, right = of TRN-2] (TRN-3) {$\dots$};
  \node [n = myblued, minimum width = 18pt, right = of TRN-3] (TRN-4) {$\mathbf{D}$};
  \node [n = mygreen, minimum width = 18pt, right = of TRN-4] (TRN-5) {$\overline{\mathbf{D}}$};
  \node [n = myred, minimum width = 36pt, right = of TRN-5] (HDR) {HDR};
  \node [n = myyellow, minimum width = 128pt, right = of HDR] (PLD) {PLD};
  \node [above = 2pt of TRN-1.north west, anchor = south west] {\scriptsize\ttfamily\textcolor{myblued}{0101010101...0101}\textcolor{mygreen}{1010}};
  \draw [decorate, decoration = {brace, mirror, raise = 2pt, amplitude = 4pt}] (TRN-1.south west) -- (TRN-5.south east) node [below = 11pt, midway, align = center] {TRN\\[-1pt]\notsotiny(7\,+\,1)\;\texttimes\;32\;=\;256\,bits};

  \node [n = mypurple, below = 8pt of HDR.south west, anchor = north west, minimum width = 25pt] (MCS) {MCS};
  \node [n = myblues, minimum width = 50pt, right = of MCS] (length) {\smash{Length}\vphantom{MCS}};
  \node [n = gray, minimum width = 75pt, right = of length] (reserved) {Reserved};
  \draw [dl] (HDR.south west) -- (MCS.north west);
  \draw [dl] (HDR.south east) -- (reserved.north east);
  \node [below = 2pt of MCS, font = \sffamily\notsotiny] {8\,bits};
  \node [below = 2pt of length, font = \sffamily\notsotiny] {16\,bits};
  \node [below = 2pt of reserved, font = \sffamily\notsotiny] {40\,bits};
\end{tikzpicture}%