### 2024-01-02 latex/tikz/transceiver_design.tex
Report in Progress

\begin{tikzpicture}[
  , thick
  , n/.style = {
    , draw
    , fill = myblued!15
    , align = center
  }
  , font = \small\sffamily
]
  \node [n] {RF Data\\Converter};
\end{tikzpicture}%