### 2024-01-04 latex/tikz/costas_loop.tex
Costas Loop TikZ Drawing

\begin{tikzpicture}[
  , thick
  , font = \small\sffamily
  , node distance = 8mm and 5mm
  , 1bit/.style = {
    , semithick
  }
  , font xs/.style = {
    , font = \notsotiny\sffamily
  }
  , n/.style = {
    , draw
    , minimum height = 5mm
    , fill = #1!15
  }
  , n/.default = myblued
  , LPF/.style = {
    , n = myyellow
    , minimum width = 6mm
  }
  , arr/.style = {
    , -latex
  }
  , arr r/.style = {
    , latex-
  }
  , port/.style = {
    , n = gray
    , signal
    , signal to = east
    , minimum height = 2mm
  }
  , port label/.style = {
    , font xs
    , align = center
    , below = 1mm
  }
  , mul add base/.style = {
    , draw = black
    , circle
    , minimum width = 3mm
    , fill = none
  }
  , add/.style = {
    , mul add base
    , append after command = {
      \pgfextra
        \draw [-, black] (\tikzlastnode.north) -- (\tikzlastnode.south);
        \draw [-, black] (\tikzlastnode.east)  -- (\tikzlastnode.west);
      \endpgfextra
    }
  }
  , sub/.style = {
    , mul add base
    , append after command = {
      \pgfextra
        \draw [-, black] (\tikzlastnode.north) -- (\tikzlastnode.south);
        \draw [-, black] (\tikzlastnode.east)  -- (\tikzlastnode.west);
        \node [above left = 0mm and 0mm of \tikzlastnode.north, inner sep = 0] {\notsotiny\resizebox*{\widthof{+}}{\heightof{-}}{-}};
        \node [above right = 0mm and 0mm of \tikzlastnode.east, inner sep = 0] {\notsotiny+};
      \endpgfextra
    }
  }
  , mul/.style = {
    , mul add base
    , append after command = {
      \pgfextra
        \draw [-, black] (\tikzlastnode.north west) -- (\tikzlastnode.south east);
        \draw [-, black] (\tikzlastnode.north east) -- (\tikzlastnode.south west);
      \endpgfextra
    }
  }
  , null node/.style = {
    , inner sep = 0
    , outer sep = 0
    , minimum width = 0
    , 
  }
  , joint/.style = {
    , circle
    , minimum height = 0.8mm
    , inner sep = 0
    , fill = #1
    , draw = #1
  }
  , joint/.default = black
  , mux/.style = {
    , n = mygreen
    , trapezium
    , trapezium angle = 60
    , shape border rotate = 90
    , minimum height = 3.8mm
    , anchor = south
  }
]
  \node (nco) [n] {NCO};
  \node (loop-filter) [LPF, right = of nco] {};

  \node (lpf-I) [LPF, above = of loop-filter] {};
  \node (lpf-Q) [LPF, below = of loop-filter] {};

  \node (mul-I) [mul, xshift = -7mm] at (nco |- lpf-I) {};
  \node (mul-Q) [mul, xshift = -7mm] at (nco |- lpf-Q) {};

  \node (in-j) [joint, left = 8mm of nco] {};

  \draw [arr] (loop-filter) -- (nco);
  \draw [arr] ([yshift = +1mm]nco.west) -| (mul-I) node [midway, right, yshift = +7mm] {$\cos$};
  \draw [arr] ([yshift = -1mm]nco.west) -| (mul-Q) node [midway, right, yshift = -7mm] {$\sin$};

  % input port
  \draw (in-j) -- ++ (-6mm, 0) node [port] {}
                               node (in-psk) [port label] {PSK\\strm.};

  % mux
  \draw [arr r] (loop-filter.east) -- ++ (5mm, 0) node (mux) [mux, right] {}
                                                  node [font xs, above right = .5mm and 1.2mm, inner sep = 0] {0}
                                                  node [font xs, below right = .5mm and 1.2mm, inner sep = 0] {1};
  \draw [1bit, shorten < = -.1mm] (mux.north)
    to [jump, jump pos = .595, |u-, jump swap] ([yshift = 5mm] in-psk |- mul-I.north) node [port] {}
                                                                                      node [port label] {is\_bpsk};

  % error output
  \node [sub, right = of mux.east, yshift = +2.2mm] (sub-QPSK) {};
  \draw [arr] (sub-QPSK) -- (sub-QPSK.west -| mux.east);
  \node [mul, right = of mux.east, yshift = -2.2mm] (mul-BPSK) {};
  \draw [arr] (mul-BPSK) -- (mul-BPSK.west -| mux.east);

  % error detector
  \node (err-detect-ctrl) [n = myred, align = center, anchor = west] at ([xshift = 10mm] mul-BPSK.east |- loop-filter.east) {Error\\Detector\\Control};
  \draw (sub-QPSK) -- (sub-QPSK -| err-detect-ctrl.west) node [above left, font xs, inner sep = .5mm, align = center] {\hphantom{Q}\\I};
  \draw (mul-BPSK) -- (mul-BPSK -| err-detect-ctrl.west) node [above left, font xs, inner sep = .5mm] {Q};
  \draw ([xshift = -3.5mm] sub-QPSK-| err-detect-ctrl.west) node [joint] {} to[jump, jump pos = .55] ++ (0, -8mm) -| (mul-BPSK);
  \draw ([xshift = -6.5mm] mul-BPSK-| err-detect-ctrl.west) node [joint] {} to[jump, jump pos = .55, jump swap] ++ (0, +8mm) -| (sub-QPSK);
  \node (err-detect) [fit = (err-detect-ctrl) (mux), draw, dashed, myred] {};
  \node [myred, anchor = north west, inner xsep = 0] at (err-detect.south west) {Error Detector};

  % loop filter tooltip
  \draw [thin, shorten < = .5mm] (loop-filter.south west) -- ++ (-2mm, -2mm)
                                                          -- ++ (-2mm, 0) node [left, font xs, inner sep = .5mm] {Loop Filter};
  \foreach \x in {I, Q} {
    \draw [arr] (mul-\x) -- (lpf-\x);
    \draw [arr] (in-j) |- (mul-\x);
    \draw [arr] (lpf-\x) -| (err-detect-ctrl) node (out-\x) [midway, joint] {};
    \draw (out-\x) -- ++ (5mm, 0) node [port, right, label={right:\x}] {};
  }

  % LPF: draw symbol and add label
  \foreach \x in {loop-filter, lpf-I, lpf-Q} {
    \draw ([xshift = -2.2mm, yshift = 1.5mm]\x.center) -- ++ (1.7mm, 0)
                                                       -- ++ (1mm, -3mm)
                                                       -- ++ (1.7mm, 0);
    \node [below = .5mm of \x, font xs, inner ysep = 0] {LPF};
  }
\end{tikzpicture}%