### 2024-01-08 latex/_tikz_sim_MIX.tex

\documentclass[journal,twoside]{IEEEtran}
\usepackage{preamble}

\pagestyle{empty}

\begin{document}
\begin{figure*}
  \setlength{\tabcolsep}{2pt}
  \sffamily\notsotiny
  \newcommand{\Signal}{DAC_vld}
  \newcommand{\Color}{mygreen}
  \begin{tabularx}{\linewidth}{@{}lX@{}}
    DAC\_vld & \input{tikz/sim/MIX} \\
    Rx\_tdata[1] & \renewcommand{\Signal}{Rx_tdata_1}\renewcommand{\Color}{myblued}\input{tikz/sim/MIX} \\
    Rx\_tdata[0] & \renewcommand{\Signal}{Rx_tdata_0}\renewcommand{\Color}{myblued}\input{tikz/sim/MIX} \\
    Rx\_tvalid & \renewcommand{\Signal}{Rx_tvalid}\renewcommand{\Color}{myblued}\input{tikz/sim/MIX} \\
    Rx\_tlast & \renewcommand{\Signal}{Rx_tlast}\renewcommand{\Color}{myblued}\input{tikz/sim/MIX} \\
    Rx\_tuser & \renewcommand{\Signal}{Rx_tuser}\renewcommand{\Color}{myblued}\input{tikz/sim/MIX} \\
    Tx\_1bit & \renewcommand{\Signal}{Tx_1bit}\renewcommand{\Color}{myred}\input{tikz/sim/MIX} \\
    Tx\_vld & \renewcommand{\Signal}{Tx_vld}\renewcommand{\Color}{myred}\input{tikz/sim/MIX} \\
    Rx\_1bit & \renewcommand{\Signal}{Rx_1bit}\renewcommand{\Color}{myblued}\input{tikz/sim/MIX} \\
  \end{tabularx}
  % Tx_1bit,Tx_vld,Rx_1bit,Rx_tdata_1,Rx_tdata_0,Rx_tvalid,Rx_tlast,Rx_tuser
\end{figure*}
\end{document}