### 2024-01-08 latex/_tikz_sim_Tx_QPSK.tex

\documentclass[journal,twoside]{IEEEtran}
\usepackage{preamble}

\pagestyle{empty}

\begin{document}
\begin{figure}
  \sffamily\notsotiny
  \begin{tabularx}{\linewidth}{@{}lX@{}}
    clk\_1M024 & \input{tikz/sim/Tx_QPSK_clk_1M} \\
    clk\_16M384 & \input{tikz/sim/Tx_QPSK_clk_16M} \\
    DAC\_bits[1:0] & \input{tikz/sim/Tx_QPSK_bits} \\
    DAC\_bits[1] & \input{tikz/sim/Tx_QPSK_bits_1} \\
    DAC\_bits[0] & \input{tikz/sim/Tx_QPSK_bits_0} \\
    DAC\_I[11:0] & \input{tikz/sim/Tx_QPSK_DAC_I} \\
    DAC\_Q[11:0] & \input{tikz/sim/Tx_QPSK_DAC_Q} \\
  \end{tabularx}%
\end{figure}
\end{document}