### 2024-01-08 latex/_tikz_ila_BPSK.tex
Update Paper

\documentclass[journal,twoside]{IEEEtran}
\usepackage{preamble}

\pagestyle{empty}

\begin{document}
\begin{figure*}
  \setlength{\tabcolsep}{2pt}
  \sffamily\notsotiny
  \newcommand\Color{mygreen}
  \newcommand\Modulation{BPSK}%
  \newcommand\isBPSK1%
  \begin{tabularx}{\linewidth}{@{}lX@{}}
    DAC\_vld & \input{tikz/ila/DAC_vld} \\
    DAC\_I[11:0] & \input{tikz/ila/DAC_I} \\
    DAC\_Q[11:0] & \input{tikz/ila/DAC_Q} \\
    ADC\_I[11:0] & \input{tikz/ila/ADC_I} \\
    ADC\_Q[11:0] & \input{tikz/ila/ADC_Q} \\
    I\_16M[15:0] & \input{tikz/ila/I_16M} \\
    Q\_16M[15:0] & \input{tikz/ila/Q_16M} \\
    clk\_1M\_out & \input{tikz/ila/clk_1M_out} \\
    I\_1M[15:0] & \input{tikz/ila/I_1M} \\
    Q\_1M[15:0] & \input{tikz/ila/Q_1M} \\
    DAC\_bits[1:0] & \input{tikz/ila/DAC_bits} \\
    DAC\_bits[1] & \input{tikz/ila/DAC_bits_1} \\
    \textcolor{gray}{DAC\_bits[0]} & \input{tikz/ila/DAC_bits_0} \\
    Rx\_data\_tdata[1:0] & \input{tikz/ila/Rx_data_tdata} \\
    Rx\_data\_tdata[1] & \input{tikz/ila/Rx_data_tdata_1} \\
    \textcolor{gray}{Rx\_data\_tdata[0]} & \input{tikz/ila/Rx_data_tdata_0} \\
    Rx\_data\_tvalid & \input{tikz/ila/Rx_data_tvalid} \\
    Rx\_data\_tlast & \input{tikz/ila/Rx_data_tlast} \\
    Rx\_data\_tuser & \input{tikz/ila/Rx_data_tuser} \\
    Tx\_1bit & \input{tikz/ila/Tx_1bit} \\
    Tx\_vld & \input{tikz/ila/Tx_vld} \\
    Rx\_1bit & \input{tikz/ila/Rx_1bit} \\
    Rx\_vld & \input{tikz/ila/Rx_vld} \\
  \end{tabularx}%
\end{figure*}
\end{document}